Put Title here

### 2018-09-26 General WLD combinatorics.tex
hopefully clarified proof of lem 2.9

\begin{lem}\label{tausimpleplanarlem}
If $W = (\cP, n)$ is an admissible Wilson loop diagram, then $\tau(W)$ is a simple planar graph whose outer face is a cycle. It is embedded such that the vertices all lie on this infinite face. These vertices are cyclically ordered, with a choice of first vertex giving it an additional compatible linear order.
\end{lem}

@@

Let $T(W)$ be the dual graph of $\tau(W)$ with the vertex corresponding to the infinite face split. In other words, place a vertex on each finite face of $\tau(W)$. These vertices are connected if there is a chord edge separating the edges. Furthermore, if the face is bounded by an edge of the polygonal cycle of $\tau(W)$, the corresponding vertex gets a leaf edge for each such boundary. Since $\tau(W)$ is a simple planar graph by Lemma \ref{tausimpleplanarlem}, $T(W)$ is an uniquely defined graph.

We claim that $T(W)$ is a tree. Since $\tau(W)$ is a planar graph with $k$ chord edges and no internal vertices, there are $k+1$ internal faces of $\tau(W)$. Each of these internal faces corresponds to a vertex of $T(W)$, and there are $n$ vertices of $T(W)$ from the splitting of the dual graph, so $T(W)$ has $n+k+1$ vertices in total. On the other hand, $T(W)$ has $n+k$ edges, corresponding to the $n +k$ edges of $\tau(W)$. Therefore, $T(W)$ is a tree.

By construction, each face of $\tau(W)$ is a polygon. Therefore, each vertex of $T(W)$ is at least 3 valent (one for each edge of the polygon). \hlfix{If split each vertex of $T(W)$ if it is strictly greater than $3$ valent.}{???} (In other words, we fail to split exactly when the face is a triangle). The connected components of $T(W)$ correspond to the decomposition of $\tau(W)$ into maximal triangulated pieces and edges originally in the polygon of $\tau(W)$.  Let $f$ be the forest thus obtained. The vertices of $f$ either have valence 1, (if they correspond leaf vertices) or 3,  (if they correspond to triangulated faces). Subtrees with no trivalent vertices correspond to either edges in the polygon of $\tau(W)$, if they were originally leaves of $T(W)$, or to maximal trivial triangulations. Splitting at all the faces that are not triangles ensures maximality of the decomposition. If the splitting were not maximal, then one could add a triangle to a connected component of the splitting, but this would imply that that splitting happened at a valence $3$ vertex.

>>>>

\begin{lem}\label{tausimpleplanarlem}
If $W = (\cP, n)$ is an admissible Wilson loop diagram, then $\tau(W)$ is a simple planar graph whose outer face is a cycle. It is embedded such that the vertices all lie on this infinite face\footnote{That is, it is an \emph{outerplanar} graph}. These vertices are cyclically ordered, with a choice of first vertex giving it an additional compatible linear order.
\end{lem}

@@

By \emph{splitting} a vertex $v$ we will mean replacing $v$ by new vertice $v_1, v_2,\ldots, v_{\text{deg}(v)}$ such that each $v_i$ as exactly one neighbour and the union of the neighbourhoods of the $v_i$ is the neighbourhood of $v$.

Let $T(W)$ be the dual graph of $\tau(W)$ with the vertex corresponding to the infinite face split.
%In other words, place a vertex on each finite face of $\tau(W)$. These vertices are connected if there is a chord edge separating the edges. Furthermore, if the face is bounded by an edge of the polygonal cycle of $\tau(W)$, the corresponding vertex gets a leaf edge for each such boundary.
Since $\tau(W)$ is an embedded graph (with a fixed distinguished embedding) by Lemma \ref{tausimpleplanarlem}, $T(W)$ is an uniquely defined graph.\note{The fixed embedding is important, because otherwise you only have a unique planar dual if the graph is 3-connected which this one isn't.  I'm just noting that here so that everyone knows why I added the part about the embedding and then we can take this comment out.}

Furthermore $T(W)$ is a tree because it is connected, has $n+k+1$ vertices ($k+1$ from the internal faces of $\tau(W)$ and $n$ from the outer face) and $n+k$ edges (since $\tau(W)$ has $n+k$ edges).  Additionally, since $\tau(W)$ is simple, $T(W)$ has no vertices of degree $2$.

%We claim that $T(W)$ is a tree. Since $\tau(W)$ is a planar graph with $k$ chord edges and no internal vertices, there are $k+1$ internal faces of $\tau(W)$. Each of these internal faces corresponds to a vertex of $T(W)$, and there are $n$ vertices of $T(W)$ from the splitting of the dual graph, so $T(W)$ has $n+k+1$ vertices in total. On the other hand, $T(W)$ has $n+k$ edges, corresponding to the $n +k$ edges of $\tau(W)$. Therefore, $T(W)$ is a tree.

Split every vertex of $T(W)$ which has degree $>3$.
%By construction, each face of $\tau(W)$ is a polygon. Therefore, each vertex of $T(W)$ is at least 3 valent (one for each edge of the polygon). \hlfix{If split each vertex of $T(W)$ if it is strictly greater than $3$ valent.}{???} (In other words, we fail to split exactly when the face is a triangle).
The connected components of $T(W)$ correspond to the decomposition of $\tau(W)$ into maximal triangulated pieces and edges originally in the polygon of $\tau(W)$.  Let $f$ be the forest thus obtained.
The vertices of $f$ either have degree 1 or 3.
%, (if they correspond leaf vertices) or 3,  (if they correspond to triangular faces).
Trees of $f$ with no trivalent vertices correspond to either edges in the polygon of $\tau(W)$, if they were originally leaves of $T(W)$, or to maximal trivial triangulated pieces.
Splitting at all the faces that are not triangles ensures maximality of the decomposition. If the splitting were not maximal, then one could add a triangle to a connected component of the splitting, but this would imply that that splitting happened at a valence $3$ vertex.
